---
title: 'Model quality: Simple and Weighted Bayes'
author: "Klara Krøyer Fomsgaard"
date: "2024-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Setup
```{r}
set.seed(1998)
setwd('~/Exam_MSc/Laura CogMod/Portfolio3/Own_files')

pacman::p_load(extraDistr, tidyverse, truncnorm, cascsim)
pacman::p_load(tidyverse,
               here,
               posterior,
               cmdstanr,
               brms, 
               tidybayes, 
               loo, job,
               bayesplot,
               ggpubr,
               extrafont)
```

## ------- SIMPLE BAYES ------- 

### Simulate data for a single agent 
```{r}
# --- Source the function for creating data
source('./Bayes_sim.R')

# --- Simulate data for 1 agent
singleParticipant_simpleBayes <- Simple_Bayes_Agent(trials = 100, # Number of trials
                   rate_own = 0.6,# Own rate - used for simulating the initial rating for the individual (Source 1)
                   rate_group = 0.4, # Group rate - used for simulating the initial rating for the group (Source 2)
                   bias = 1) # Bias used when making the second rating

singleParticipant_simpleBayes$ID <- 2
```

```{r}
# --- Plot over simulated ratings: Simple Bayes

# First rating
sim_first_plot <- singleParticipant_simpleBayes %>% 
  ggplot()+
  geom_histogram(aes(FirstRating_s1), fill = 'darkcyan',alpha = .8) +
  theme_minimal() +
  labs(title = "First Rating", x = "Ratings", y = "Count")+
  theme(text=element_text(size=11,  family="Georgia"))

# Group rating
sim_group_plot <- singleParticipant_simpleBayes %>% 
  ggplot()+
  geom_histogram(aes(GroupRating_s2), fill = 'darkcyan',alpha = .8) +
  theme_minimal() +
  labs(title = "Group Rating", x = "Ratings", y = "Count")+
  theme(text=element_text(size=11,  family="Georgia"))

# Second rating
sim_second_plot <- singleParticipant_simpleBayes %>% 
  ggplot()+
  geom_histogram(aes(SecondRating_predicted), fill = 'darkcyan',alpha = .8) +
  theme_minimal() +
  labs(title = "Second Rating", x = "Ratings", y = "Count")+
  theme(text=element_text(size=11,  family="Georgia"))

# Collect in one figure
figure <- ggarrange(sim_first_plot, sim_second_plot, sim_group_plot,
                    labels = c("A", "B", "C"),
                    ncol = 3, nrow = 1)


figure
```

### Model fitting
#### Extract data for STAN
```{r}
# fixing it to be STAN format
simple_data_list <- list(
  trials = nrow(singleParticipant_simpleBayes),
  FirstRating = singleParticipant_simpleBayes$FirstRating_s1,
  GroupRating = singleParticipant_simpleBayes$GroupRating_s2,
  SecondRating = singleParticipant_simpleBayes$SecondRating_predicted
)
```

#### Load STAN model
```{r}
# Loading model
file_simple <- file.path("./simple_Bayes_model.stan")
simple_model <- cmdstan_model(file_simple, 
                     cpp_options = list(stan_threads = TRUE),
                     stanc_options = list("O1"))
```

#### --- Fit model to simulated data
```{r}
samples_simple <- simple_model$sample(
  data = simple_data_list,
  seed = 123,
  chains = 4,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 0,
  max_treedepth = 20,
  adapt_delta = 0.99,
)
```
#### Inspect model fit
```{r}
# Model diagnostics
samples_simple$cmdstan_diagnose()

# Trace plots
trace <- mcmc_trace(samples_simple$draws("bias"))

# Save samples from the posterior
draws_simple <- as_draws_df(samples_simple$draws()) 
head(draws_simple)

# Prior-Posterior update plot
simple_sim_pp_update <- ggplot() +
  geom_density(aes(draws_simple$bias), fill = "darkcyan", alpha = 0.5, color = "darkcyan") + ### posterior from model 
  geom_density(aes(rnorm(4000,0,2)), fill = "grey", alpha = 0.5, color = "grey") + ### just directly simulated as it isnt nested in the model 
  geom_vline(xintercept = 1)+
  xlab("Rate") +
  ylab("Density") +
  labs(title="Prior Posterior Check, Bias", subtitle = "Simple Bayes") +
  theme_minimal()+
  theme(text=element_text(size=11,  family="Georgia"))+
  xlim(-5,5)

results_plot <- ggarrange(trace, simple_sim_pp_update,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2)
```

## ------ WEIGHTED BAYES ------
This model is an extension of the simple Bayes, but it includes weights for each source.

### Simulate a weighted Bayes agent
```{r}
trials <- 100 # Number of trials
rate_group <- .4 # Group rate - used for simulating the initial rating for the group (Source 2)
rate_own <- .6 # Own rate - used for simulating the initial rating for the individual (Source 1)
bias <- 1 # Bias for the participant when making the second rating

w1 <- .6 # Weight for Source 1
w2 <- .4 # Weight for Source 2

Weighted_Bayes <- Weighted_Bayes_Agent(trials = trials,
                     rate_own = rate_own,
                     rate_group = rate_group,
                     bias = bias,
                     w1 = w1, 
                     w2 = w2)
```

```{r}
# --- Plot over simulated ratings: Weighted Bayes

# First rating
weight_first_plot <- Weighted_Bayes %>% 
  ggplot()+
  geom_histogram(aes(FirstRating_s1), fill = 'darkcyan',alpha = .8) +
  theme_minimal() +
  labs(title = "First Rating", x = "Ratings", y = "Count")+
  theme(text=element_text(size=11,  family="Georgia"))

# Group rating
weight_group_plot <- Weighted_Bayes %>% 
  ggplot()+
  geom_histogram(aes(GroupRating_s2), fill = 'darkcyan',alpha = .8) +
  theme_minimal() +
  labs(title = "Group Rating", x = "Ratings", y = "Count")+
  theme(text=element_text(size=11,  family="Georgia"))

# Second rating
weight_second_plot <- Weighted_Bayes %>% 
  ggplot()+
  geom_histogram(aes(SecondRating_predicted), fill = 'darkcyan',alpha = .8) +
  theme_minimal() +
  labs(title = "Second Rating", x = "Ratings", y = "Count")+
  theme(text=element_text(size=11,  family="Georgia"))

# Collect in one figure
figure_weight <- ggarrange(weight_first_plot, weight_group_plot, weight_second_plot,
                    labels = c("A", "B", "C"),
                    ncol = 3, nrow = 1)


figure_weight
```


### Fit model: Weighted Bayes to simulated data

#### Extract data for STAN
```{r}
# fixing it to be stan format
weighted_data_list <- list(
  trials = nrow(Weighted_Bayes),
  FirstRating = Weighted_Bayes$FirstRating_s1,
  GroupRating = Weighted_Bayes$GroupRating_s2,
  SecondRating = Weighted_Bayes$SecondRating_predicted
)
```

#### Load STAN model
```{r}
# Loading model
file_weighted <- file.path("./weighted_Bayes_model.stan")
model_weighted <- cmdstan_model(file_weighted, 
                     cpp_options = list(stan_threads = TRUE),
                     stanc_options = list("O1"))
```

```{r}
# Fit model
samples_weighted <- model_weighted$sample(
  data = weighted_data_list,
  seed = 123,
  chains = 4,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 0,
  max_treedepth = 20,
  adapt_delta = 0.99,
)
```

```{r, warning = FALSE}
# Save samples from the posterior
draws_weighted <- as_draws_df(samples_weighted$draws()) 
head(draws_weighted)

# Trace plots
trace_bias <- mcmc_trace(samples_weighted$draws("bias"))
trace_w1 <- mcmc_trace(samples_weighted$draws("w1"))
trace_w2 <- mcmc_trace(samples_weighted$draws("w2"))

ggarrange(trace_bias, trace_w1, trace_w2,
                    labels = c("A", "B", "C"),
                    ncol = 1, nrow = 3)

# --- Plots: Prior Posterior update checks
weighted_pp_update_bias <- ggplot() +
  geom_density(aes(draws_weighted$bias), fill = "darkcyan", alpha = 0.5, color = "darkcyan") + ### posterior from model 
  geom_density(aes(rnorm(4000,0,2)), fill = "grey", alpha = 0.5, color = "grey") + ### just directly simulated as it isnt nested in the model 
  geom_vline(xintercept = 1)+
  xlab("Rate") +
  ylab("Density") +
  labs(title="Prior Posterior Check, Bias", subtitle = "Weighted Bayes") +
  theme_minimal()+
  theme(text=element_text(size=11,  family="Georgia"))+
  xlim(-5,5)

# Prior-Posterior update plot - W1
weighted_pp_update_w1 <- ggplot() +
  geom_density(aes(draws_weighted$w1), fill = "darkcyan", alpha = 0.5, color = "darkcyan") + ### posterior from model 
  geom_density(aes(rbeta(4000,2,2)), fill = "grey", alpha = 0.5, color = "grey") + ### just directly simulated as it isnt nested in the model 
  geom_vline(xintercept = .6)+
  xlab("Rate") +
  ylab("Density") +
  labs(title="Prior Posterior Check, W1", subtitle = "Weighted Bayes") +
  theme_minimal()+
  theme(text=element_text(size=11,  family="Georgia"))+
  xlim(0,1)

# Prior-Posterior update plot - W2
weighted_pp_update_w2 <- ggplot() +
  geom_density(aes(draws_weighted$w2), fill = "darkcyan", alpha = 0.5, color = "darkcyan") + ### posterior from model 
  geom_density(aes(rbeta(4000,2,2)), fill = "grey", alpha = 0.5, color = "grey") + ### just directly simulated as it isnt nested in the model 
  geom_vline(xintercept = .4)+
  xlab("Rate") +
  ylab("Density") +
  labs(title="Prior Posterior Check, W2", subtitle = "Weighted Bayes") +
  theme_minimal()+
  theme(text=element_text(size=11,  family="Georgia"))+
  xlim(0,1)

weighted_pp <- ggarrange(weighted_pp_update_bias, weighted_pp_update_w1, weighted_pp_update_w2,
                    labels = c("A", "B", "C"),
                    ncol = 3, nrow = 1)

```
