---
title: "Installing and using R environment on ucloud"
author: "Malte Lau Petersen"
date: "2024-02-13"
output: html_document
---

# R environments

Just like python, environments are a mess in R but you can sort of get an OK solution if you're lucky that your use case lines up with the developers of the tools. I use `renv` and think about 3 distinct actions that are possible.

1. Initialize the environment and save it somewhere it's not going to get deleted when you shut down the ucloud instance. This can be in `/work/Home/` or in a project folder, but you probably don't want to commit this to git.
2. Install packages into the environment
3. Activate the environment to use the packages

In order to set up your environment, do step 1 and 2 (which are both in the `install.R` script below). This should only be needed once. When you have your environment set up and want to run something in a fresh ucloud container, do step 3 (and optionally step 2) which is in the `activate.R` script below. The easiest way to do that is to just add `source("/path/to/activate.R")` at the top of your R scripts.

# install.R
```{r install, eval=FALSE}
workdir = "/work/Home/R"

setwd(workdir)
Sys.setenv(RENV_PATHS_CACHE = paste0(workdir, '/renv/cache'))
options("renv.consent" = TRUE)

renv::init() ## takes several minutes
install.packages(c("tidyverse", "brms"))
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
cmdstanr::install_cmdstan(dir = workdir)
renv::snapshot(prompt = FALSE)

```

# activate.R
```{r activate, eval=FALSE}
workdir = '/work/Home/R'

Sys.setenv(RENV_PATHS_CACHE = paste0(workdir, '/renv/cache'))
options("renv.consent" = TRUE)
renv::load(project = workdir)
cmdstanr::set_cmdstan_path(paste0(workdir, '/cmdstan-2.34.1')) # adapt for newer versions of cmdstanr


Sys.setenv(
  MAKEFLAGS = paste0(
    "-j", 
    parallel::detectCores(logical = FALSE)
  ))

```

# Installing new packages
if you end up installing new packages, finish with
```{r newpackages, eval=FALSE}
install.packages(...)
renv::snapshot()
```

